---
title: "Urogynecology and Reconstructive Pelvic Surgery (URPS) Literature, 1975–2024"
subtitle: "A longitudinal bibliometric analysis"
author: "Tyler Muffly, MD"
date: "`r format(Sys.Date(), '%d %B %Y')`"
params:
  results_rds: "output/pipeline_result.rds"
  refresh: FALSE
  year_start: 1975
  year_end: 2024
  focal: "URPS"
  top_n: 10
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    fig_caption: true
    df_print: kable
    theme: readable
  word_document:
    toc: true
    fig_caption: true
    reference_docx: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  fig.align  = "center",
  fig.width  = 7.5,
  fig.height = 5,
  dpi        = 200,
  out.width  = "100%"
)

# knitr starts in the Rmd's folder; the pipeline and its output/ live one
# level up. Every chunk after this one therefore runs at the repo root, so
# all paths below (including params$results_rds) are repo-root-relative.
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load-pipeline, include=FALSE}
suppressWarnings(suppressMessages(
  sys.source("R/fpmrs_bibliometrics_pipeline.R", envir = globalenv())
))
source("manuscript/_helpers.R", local = knitr::knit_global())
```

```{r analysis, include=FALSE}
res <- load_or_run_analysis(
  results_rds = params$results_rds,
  refresh     = params$refresh,
  year_start  = as.integer(params$year_start),
  year_end    = as.integer(params$year_end)
)

focal_row <- build_focal_row(res, focal = params$focal)

# Equity is recomputed here rather than read from the cached object: a cache
# written before the ISO-2 income-tier fix carries the old, wrong shares.
equity <- .compute_equity_metrics(res$bibliography, verbose = FALSE)

abstract <- generate_abstract_results_text(
  comparison_summary_table  = focal_row,
  focal_subspecialty        = params$focal,
  focal_mk_trend            = res$mk_trend,
  focal_equity_metrics      = equity,
  focal_institution_metrics = res$institution_metrics,
  focal_keyword_trends      = res$keyword_trends,
  year_start                = as.integer(params$year_start),
  year_end                  = as.integer(params$year_end),
  verbose                   = FALSE
)

has_comparators <- comparative_sentences_available(focal_row)

n_docs  <- nrow(res$bibliography)
n_years <- length(unique(res$annual_trends$publication_year))
```

```{r caveat, results="asis"}
if (!has_comparators) {
  cat(
    "> **Single-corpus run.** Only the ", params$focal, " corpus is loaded, so ",
    "every rank is 1 by construction and the comparative sentences ",
    "(cohort ranking, urology contrast, evidence-quality benchmarking) are ",
    "either absent or trivially true. Run the pipeline for the comparator ",
    "subspecialties and pass the results through ",
    "`run_subspecialty_comparison()` to populate them.\n\n",
    sep = ""
  )
}
```

# Abstract

**`r abstract$title`**

**Introduction.** `r abstract$introduction`

**Methods.** `r abstract$methods`

**Results.** `r paste(na.omit(unlist(abstract[c("corpus","citations","growth","mk_trend","geography","institution","output","keywords","disruption","evidence_quality","equity","collaboration","urology")])), collapse = " ")`

**Conclusion.** `r abstract$conclusion`

```{r word-count, results="asis"}
body <- paste(
  na.omit(unlist(abstract[setdiff(names(abstract), c("title", "formatted"))])),
  collapse = " "
)
cat(sprintf(
  "\n*Abstract body: %d words.*\n",
  lengths(gregexpr("\\S+", body))
))
```

# Methods

`r abstract$methods`

## Search strategy

The full PubMed query is reproduced in the [Reproducibility] appendix. Its
SHA-256 hash is recorded with every run so a published figure can be traced
to the exact search that produced it.

## Record flow

```{r tbl-consort}
consort <- consort_table(file.path("output", "consort_flow.json"))
if (!is.null(consort)) {
  manuscript_table(consort, "Table 1. CONSORT-style record flow.")
}
```

# Results

## Publication volume

`r abstract$corpus` `r abstract$growth`

`r abstract$mk_trend`

```{r fig-annual, fig.cap="Figure 1. Annual URPS publication volume, 1975-2024, with LOESS trend and 95% confidence interval."}
plot_annual_publications(
  res$annual_trends,
  year_start = as.integer(params$year_start),
  year_end   = as.integer(params$year_end),
  verbose    = FALSE
)
```

## Citation impact

`r abstract$citations`

```{r fig-citations, fig.height=6, fig.cap="Figure 2. Citation impact over time. Both panels index citations to the year of publication; the post-2000s decline is a censoring artifact of shorter citation windows, not falling impact."}
plot_citation_trends(
  res$annual_trends,
  year_start = as.integer(params$year_start),
  year_end   = as.integer(params$year_end),
  verbose    = FALSE
)
```

## Journals

`r abstract$output`

```{r fig-journals, fig.height=6, fig.cap="Figure 3. Journal dominance in URPS literature over time (top 8 by total volume)."}
plot_journal_trends(res$journal_trends, top_n_journals = 8L, verbose = FALSE)
```

```{r tbl-journals}
manuscript_table(
  journal_table(res, top_n = params$top_n),
  sprintf("Table 2. Top %d journals by total citations.", params$top_n)
)
```

## Authors

```{r tbl-authors}
manuscript_table(
  author_table(res, top_n = params$top_n),
  sprintf("Table 3. Top %d authors by publication count.", params$top_n)
)
```

## Geography and equity

`r abstract$geography`

`r abstract$equity`

```{r fig-countries, fig.height=6, fig.cap="Figure 4. Geographic distribution of URPS publications (top 10 first-author countries)."}
plot_country_contributions(res$country_trends, top_n_countries = 10L, verbose = FALSE)
```

```{r tbl-countries}
manuscript_table(
  country_table(res, top_n = 15L),
  "Table 4. Top 15 first-author countries with World Bank income tier."
)
```

```{r tbl-equity}
manuscript_table(
  equity_table(equity),
  "Table 5. Distribution of publications by World Bank income group."
)
```

## Research themes

`r abstract$keywords`

```{r fig-keywords, fig.height=6, fig.cap="Figure 5. Keyword evolution over time (top 20 author keywords). Author keywords are sparsely indexed in PubMed before ~2000, so early blank cells reflect missing metadata rather than absent topics."}
plot_keyword_evolution(res$keyword_trends, top_n_keywords = 20L, verbose = FALSE)
```

## Collaboration

`r abstract$collaboration`

# Conclusion

`r abstract$conclusion`

# Reproducibility

```{r repro, results="asis"}
qh <- file.path("output", "query_hash.txt")
if (file.exists(qh)) {
  cat("**Query hash**\n\n```\n", paste(readLines(qh), collapse = "\n"), "\n```\n\n", sep = "")
}
cat("**PubMed query**\n\n```\n",
    paste(strwrap(get_subspecialty_pubmed_query("fpmrs"), width = 76), collapse = "\n"),
    "\n```\n\n", sep = "")
cat(sprintf(
  "Corpus: %s records across %d publication years (%d-%d).\n\n",
  format(n_docs, big.mark = ","), n_years,
  as.integer(params$year_start), as.integer(params$year_end)
))
```

```{r session}
si <- sessionInfo()
cat("R version:", si$R.version$version.string, "\n")
cat("Platform :", si$platform, "\n")
key <- c("dplyr", "ggplot2", "stringr", "tibble", "purrr", "scales", "jsonlite")
vers <- vapply(key, function(p) {
  if (requireNamespace(p, quietly = TRUE))
    as.character(utils::packageVersion(p)) else NA_character_
}, character(1L))
cat("\nKey packages:\n")
print(data.frame(package = key, version = unname(vers), row.names = NULL))
```
