---
title: "Fifty Years of Urogynecology and Reconstructive Pelvic Surgery Research: A Bibliometric Analysis of 49,355 Publications, 1975-2024"
author:
  - Tyler Muffly, MD
date: "`r format(Sys.Date(), '%d %B %Y')`"
params:
  results_rds: "output/pipeline_result.rds"
bibliography: references.bib
csl: urogynecology.csl
link-citations: false
output:
  word_document:
    toc: false
    fig_caption: true
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 5, dpi = 300, fig.align = "center"
)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r data, include=FALSE}
suppressWarnings(suppressMessages(
  sys.source("R/fpmrs_bibliometrics_pipeline.R", envir = globalenv())
))
source("manuscript/_helpers.R", local = knitr::knit_global())

res <- readRDS(params$results_rds)
at  <- res$annual_trends
eq  <- .compute_equity_metrics(res$bibliography, verbose = FALSE)
cm  <- res$citation_metrics
mk  <- res$mk_trend
gm  <- res$growth_metrics

n_docs     <- nrow(res$bibliography)
n_cit      <- sum(at$total_citations, na.rm = TRUE)
mean_cit   <- round(n_cit / n_docs, 1)
first_yr   <- at$publication_year[1]
first_n    <- at$publication_count[1]
last_yr    <- utils::tail(at$publication_year, 1)
last_n     <- utils::tail(at$publication_count, 1)
fold       <- round(last_n / first_n, 1)

co <- res$country_summary
co <- co[!is.na(co$country) & !toupper(co$country) %in% c("NA", "NULL") &
           nzchar(co$country), ]
n_countries <- nrow(co)
top_ctry_pct <- round(co$pct_of_total[[1L]], 1)
lmic_pct <- round(eq$pct_lower_mid + eq$pct_low_income, 1)

fmt <- function(x) format(x, big.mark = ",", trim = TRUE)
pv  <- function(p) paste0("P", .fmt_pvalue(p))

consort <- consort_table("output/consort_flow.json")
n_retrieved <- consort$Records[1]
n_excl_year <- consort$Records[2]
n_excl_lang <- consort$Records[4]
```

# Structured Abstract

**Importance.** Urogynecology and Reconstructive Pelvic Surgery (URPS) treats
pelvic floor disorders affecting roughly one in four women,[@nygaard2008] yet
the field's research output has never been characterized across its full
modern history.

**Objectives.** To quantify the volume, growth, citation impact, and global
authorship of URPS literature from `r first_yr` to `r last_yr`.

**Study Design.** We searched PubMed using Medical Subject Headings and
title/abstract terms spanning both FPMRS and URPS nomenclature, plus the
subspecialty journal under both its names. Records were linked to OpenAlex by
PubMed identifier for citation counts and first-author country.[@priem2022]
Analyses used the bibliometrix R package.[@aria2017] Trend was tested by
Mann-Kendall and segmented regression; authorship equity was summarized by
World Bank income group and the Herfindahl-Hirschman Index (HHI).

**Results.** `r fmt(n_retrieved)` records were retrieved and
`r fmt(n_docs)` retained after year and language filtering. Annual output rose
from `r first_n` publications in `r first_yr` to `r fmt(last_n)` in
`r last_yr`, a `r fold`-fold increase (compound annual growth rate
`r gm$cagr_pct`%). Growth was monotonic (Mann-Kendall tau=`r round(mk$mk_tau,3)`,
`r pv(mk$mk_p)`) with a structural inflection in `r mk$breakpoint_year`, after
which output accelerated from `r round(mk$seg1_slope,1)` to
`r round(mk$seg2_slope,1)` publications per year. The corpus accrued
`r fmt(n_cit)` citations (mean `r mean_cit` per paper; h-index
`r cm$h_index`; `r round(cm$pct_uncited,1)`% uncited). Authorship spanned
`r n_countries` countries but was concentrated in high-income settings
(`r eq$pct_high_income`%), with `r lmic_pct`% from lower-middle- or
low-income countries.

**Conclusions.** URPS research has grown steadily for five decades and
accelerated sharply after `r mk$breakpoint_year`. Output remains dominated by
high-income countries, identifying a clear target for international research
partnership.

```{r abstract-count, results="asis"}
# The abstract above is written by hand; this reports its length so the
# 250-word journal limit can be checked at every knit rather than eyeballed.
rmd <- readLines("manuscript/urps_journal_manuscript.Rmd", warn = FALSE)
beg <- grep("^# Structured Abstract", rmd)
end <- grep("^```\\{r abstract-count", rmd)
txt <- paste(rmd[(beg + 1):(end - 1)], collapse = " ")
txt <- gsub("`r [^`]*`", "XXXX", txt)          # inline code -> one token
txt <- gsub("\\[@[^]]*\\]", "", txt)           # citations are not words
txt <- gsub("[*_#]", "", txt)
n_words <- lengths(gregexpr("\\S+", trimws(txt)))
cat(sprintf("\n*Abstract length: %d words (journal limit 250).*\n", n_words))
```

# Introduction

Pelvic floor disorders -- urinary incontinence, pelvic organ prolapse, and
fecal incontinence -- affect approximately one quarter of adult women in the
United States, and their prevalence rises steeply with age.[@nygaard2008] A
woman's lifetime risk of surgery for prolapse or incontinence approaches 11%
by age 80.[@olsen1997] Population surveys across five countries confirm that
lower urinary tract symptoms are common and burdensome
internationally.[@irwin2006]

The subspecialty addressing these conditions has undergone rapid institutional
change. Standardized terminology for prolapse quantification arrived in
1996,[@bump1996] followed by a joint International Urogynecological
Association / International Continence Society terminology report in
2010.[@haylen2010] Validated patient-reported instruments such as the ICIQ
followed a similar trajectory.[@avery2004] The subspecialty itself was renamed
from Female Pelvic Medicine and Reconstructive Surgery (FPMRS) to
Urogynecology and Reconstructive Pelvic Surgery (URPS), effective 2024.

Despite this maturation, the field's research output has not been described
across its full modern history. Bibliometric analysis quantifies how a
discipline's literature grows, where its influence concentrates, and who is
excluded from producing it. We therefore characterized `r fmt(n_docs)` URPS
publications spanning `r first_yr` to `r last_yr`.

# Materials and Methods

## Search strategy and data sources

We searched PubMed for records published between `r first_yr` and `r last_yr`.
The strategy combined Medical Subject Headings (pelvic organ prolapse, pelvic
floor disorders, stress and urge urinary incontinence, cystocele, rectocele,
uterine prolapse) with title/abstract terms covering both the FPMRS and URPS
names and their clinical vocabulary (sacrocolpopexy, colporrhaphy,
suburethral and pubovaginal sling, colposuspension). Because the subspecialty
journal was renamed in 2022, both *Female Pelvic Medicine and Reconstructive
Surgery* and *Urogynecology* were included as journal terms. The complete
query and its SHA-256 hash accompany the analysis code.

PubMed does not supply citation counts or structured affiliation countries.
Records were therefore linked to OpenAlex by PubMed identifier to obtain
citation counts and the first author's affiliation country.[@priem2022]
Citation and geographic analyses are restricted to records with a successful
match.

## Analysis

Analyses were performed in R[@rcore] using the bibliometrix package.[@aria2017]
Annual publication counts, citation totals, and journal, author, keyword, and
country distributions were computed from the retained corpus. Monotonic trend
was assessed by the Mann-Kendall test; a structural break in the growth
trajectory was located by segmented regression. Citation concentration was
summarized by the h-index and the Gini coefficient. Author affiliation
countries were mapped to World Bank income groups, and geographic
concentration was summarized by the Herfindahl-Hirschman Index (HHI). No
patient-level data were used and institutional review board approval was not
required.

## Record flow

```{r tbl-consort}
manuscript_table(consort, "Table 1. Record flow from search to analysis.")
```

# Results

## Publication volume and growth

Of `r fmt(n_retrieved)` records retrieved, `r fmt(n_excl_year)` fell outside
the study window and `r fmt(n_excl_lang)` were non-English, leaving
`r fmt(n_docs)` publications for analysis (Table 1).

Annual output rose from `r first_n` publications in `r first_yr` to
`r fmt(last_n)` in `r last_yr`, a `r fold`-fold increase corresponding to a
compound annual growth rate of `r gm$cagr_pct`% (Figure 1). The increase was
monotonic across the full period (Mann-Kendall tau=`r round(mk$mk_tau, 3)`,
`r pv(mk$mk_p)`). Segmented regression identified a structural inflection in
`r mk$breakpoint_year`, before which output grew by
`r round(mk$seg1_slope, 1)` publications per year and after which it grew by
`r round(mk$seg2_slope, 1)` per year -- a
`r round(mk$seg2_slope / mk$seg1_slope, 1)`-fold acceleration.

```{r fig-annual, fig.cap="Figure 1. Annual URPS publication volume, 1975-2024, with LOESS trend and 95% confidence interval."}
plot_annual_publications(at, min(at$publication_year), max(at$publication_year),
                         verbose = FALSE)
```

## Citation impact

The corpus accrued `r fmt(n_cit)` citations, a mean of `r mean_cit` per paper,
with an h-index of `r cm$h_index`. Citations were unevenly distributed (Gini
coefficient `r round(cm$citation_gini, 3)`), and
`r round(cm$pct_uncited, 1)`% of publications were never cited. Apparent
declines in per-paper citation counts for recent years reflect shorter
citation windows rather than falling impact (Figure 2).

```{r fig-citations, fig.height=6, fig.cap="Figure 2. Citation impact by year of publication. The post-2000s decline is a censoring artifact of shorter citation windows."}
plot_citation_trends(at, min(at$publication_year), max(at$publication_year),
                     verbose = FALSE)
```

## Journals and authors

```{r tbl-journals}
manuscript_table(journal_table(res, top_n = 10L),
                 "Table 2. Ten most-cited journals in the URPS corpus.")
```

```{r fig-journals, fig.height=6, fig.cap="Figure 3. Journal publication volume over time (top 8 journals)."}
plot_journal_trends(res$journal_trends, top_n_journals = 8L, verbose = FALSE)
```

Publications averaged `r round(res$authorship_metrics$mean_authors_per_paper, 1)`
authors per paper.

```{r tbl-authors}
manuscript_table(author_table(res, top_n = 10L),
                 "Table 3. Ten most prolific authors by publication count.")
```

## Geographic distribution and authorship equity

Authorship spanned `r n_countries` countries, but `r top_ctry_pct`% of
publications originated from the single most productive country (Figure 4,
Table 4). By World Bank income group, `r eq$pct_high_income`% of publications
came from high-income countries, `r eq$pct_upper_mid`% from upper-middle-income
countries, and `r lmic_pct`% from lower-middle- or low-income countries
combined (Table 5). Geographic concentration was moderate
(HHI `r round(eq$concentration_hhi, 3)`).

```{r fig-countries, fig.height=6, fig.cap="Figure 4. Geographic distribution of URPS publications by first-author country (top 10)."}
plot_country_contributions(res$country_trends, top_n_countries = 10L,
                           verbose = FALSE)
```

```{r tbl-countries}
manuscript_table(country_table(res, top_n = 10L),
                 "Table 4. Ten most productive first-author countries.")
```

```{r tbl-equity}
manuscript_table(equity_table(eq),
                 "Table 5. Publications by World Bank income group.")
```

## Research themes

```{r fig-keywords, fig.height=6, fig.cap="Figure 5. Author keyword frequency over time (top 20 keywords). PubMed indexes author keywords sparsely before approximately 2000."}
plot_keyword_evolution(res$keyword_trends, top_n_keywords = 20L, verbose = FALSE)
```

# Discussion

URPS research output grew `r fold`-fold over five decades and accelerated
sharply after `r mk$breakpoint_year`, from `r round(mk$seg1_slope, 1)` to
`r round(mk$seg2_slope, 1)` publications per year. That inflection follows the
period in which the field standardized how it measures disease: the pelvic
organ prolapse quantification system was published in 1996,[@bump1996] and
validated symptom instruments followed.[@avery2004] Shared measurement makes
studies comparable and cumulative, and the literature expanded accordingly.
Subsequent consolidation of terminology[@haylen2010] coincided with continued
growth.

The clearest actionable finding is geographic. Authorship spanned
`r n_countries` countries, yet `r eq$pct_high_income`% of publications
originated in high-income settings and only `r lmic_pct`% in lower-middle- or
low-income countries. Pelvic floor disorders are not confined to wealthy
countries; obstetric fistula and severe prolapse impose their heaviest burden
where access to surgical care is least. A literature produced almost entirely
in high-income settings risks generalizing findings from health systems
unlike those where much of the disease occurs. Targeted fellowship exchange,
co-authored multinational registries, and journal policies that support
authors from low-resource settings are concrete levers.

## Limitations

Several limitations warrant emphasis. First, the search retrieves records
containing pelvic floor terminology, and unqualified terms such as "urinary
incontinence" also capture literature outside the female pelvic floor,
including post-prostatectomy incontinence; some proportion of the corpus is
therefore outside the subspecialty's scope. Second, citation counts and
affiliation countries derive from OpenAlex records matched by PubMed
identifier, so unmatched records are absent from those analyses. Third, only
the first author's country is available, so international collaboration rates
could not be estimated and the geographic findings describe lead authorship
rather than all contributing countries. Fourth, restricting to English-language
records excludes `r fmt(n_excl_lang)` publications and may understate output
from non-anglophone countries, compounding the geographic concentration we
report. Fifth, recent publication years have shorter citation windows, so
per-paper impact for those years is censored downward. Finally, bibliometric
volume measures attention and activity, not clinical quality or patient
benefit.

## Conclusions

URPS research has grown continuously for fifty years, accelerating after
`r mk$breakpoint_year` in step with the field's adoption of standardized
measurement. The literature remains concentrated in high-income countries,
which is the most tractable target for deliberate change.

# Poster bullets

*Working notes for poster layout; delete before journal submission.*

**Background**

- Pelvic floor disorders affect ~1 in 4 US women; lifetime surgery risk ~11% by age 80.
- Subspecialty renamed FPMRS to URPS effective 2024.
- Research output across the field's full modern history had not been described.

**Methods**

- PubMed search, `r first_yr`-`r last_yr`, MeSH plus title/abstract terms covering both names.
- Linked to OpenAlex by PMID for citations and first-author country.
- `r fmt(n_retrieved)` retrieved, `r fmt(n_docs)` analyzed.

**Results**

- `r first_n` publications in `r first_yr` to `r fmt(last_n)` in `r last_yr` -- `r fold`-fold growth.
- Compound annual growth rate `r gm$cagr_pct`%.
- Monotonic increase: Mann-Kendall tau=`r round(mk$mk_tau, 3)`, `r pv(mk$mk_p)`.
- Structural inflection `r mk$breakpoint_year`: `r round(mk$seg1_slope, 1)` to `r round(mk$seg2_slope, 1)` papers/year.
- `r fmt(n_cit)` total citations; mean `r mean_cit`/paper; h-index `r cm$h_index`.
- `r round(cm$pct_uncited, 1)`% of papers never cited; citation Gini `r round(cm$citation_gini, 3)`.
- `r n_countries` countries represented; `r top_ctry_pct`% from the top country.
- `r eq$pct_high_income`% high-income vs `r lmic_pct`% lower-middle/low-income.

**Conclusions**

- Fifty years of uninterrupted growth, accelerating after `r mk$breakpoint_year`.
- Standardized measurement preceded the acceleration.
- High-income concentration is the field's clearest equity gap.

# References
